### 2010-12-03 report/report.tex
report update

\section{Introduction}
% Motivate and abstractly describe the problem you are addressing and
% how you are addressing it. What is the problem? Why is it important?
% What is your basic approach? A short discussion of how it fits into
% related work in the area is also desirable. Summarize the basic results
% and conclusions that you will present.
We are using machine learning to design a gait for a quadruped
robot. The robot has an on-board computer, drivers for the motors, and
will have camera feedback. The code so far consists of the robot
class, optimization class, parameterized model, sine model, and camera
feedback.  Both the hardware and software are progressing on schedule.
Some promising gaits have been generated, and we anticipate a final
evaluation of the robot’s gait in November.

\section{Problem definition}
% Precisely define the problem you are addressing (i.e. formally specify
% the inputs and outputs). Elaborate on why this is an interesting and
% important problem.
We are testing several different learning methods to design a gait for 
a quadraped robot from the Cornell Computational Synthesis Lab. The metric 
for evaluation of the designed gait is speed. A comparison and evaluation 
of the many different methods available for optimizing the gait of legged 
robots will be useful for future work on this challenging multidimensional 
control problem.

\section{Method}
% Describe in reasonable detail the algorithm you are using to address this 
% problem. A pseudo-code description of the algorithm you are using is 
% frequently useful. If it makes sense for your project, trace through a 
% concrete example, showing how your algorithm processes this example. The 
% example should be complex enough to illustrate all of the important aspects 
% of the problem but simple enough to be easily understood. If possible, 
% an intuitively meaningful example is better than one with meaningless 
% symbols.
We implemented and tested 8 different learning strategies:
\begin{itemize}
\item \emph{Uniform random hill climbing}: Creates a random neighbor by
  randomly choosing one parameter to adjust, and changing it completely 
  randomly in UniformStrategy(). Then evaluates this neighbor by running 
  the robot with the newly chosen parameters. If this neighbor results in a
  longer distance walked then the previous best neighbor, we save this
  neighbor as the new best neighbor. We then repeat the whole process
  using the best neighbor as a base.
\item \emph{Gaussian random hill climbing}: Creates a random neighbor by
  changing each parameter randomly based on a normal distribution. Then
  evaluates this neighbor similarly to uniform random hill climbing.
\item \emph{N-dimensional policy gradient descent}: Estimates the policy
  gradient by evaluating \emph{t} randomly generated policies
  near an initial parameter vector. Then computes the average score for
  each parameter and adjusts the base policy according to the estimated
  gradient.
\item \emph{Random}: Randomly generates policies in the range of the motion
representation.
\item \emph{Simplex (Nelder-Mead) Method}
\item \emph{Linear regression}: To initialize, chooses five random points
and trains on them using least-squares. Then in a loop, takes a fixed-size
step in the direction of the gradient, and retrains on all points so far.
\item \emph{SVM regression}
\item \emph{Evolutionary Neural Network (HyperNEAT)\cite{2}}
\end{itemize}

@@

The quadraped robot has an on-board computer running Linux. The lower
level drivers are in C and we are implementing the system in
Python. Feedback about distance travelled is provided via a Wii remote.

@@

\section{Experimental (or Theoretical) Evaluation}

\subsection{Methodology}
% What are the criteria you are using to evaluate your method? What
% specific hypotheses does your experiment test? Describe the 
% experimental methodology that you used. What are the dependent and 
% independent variables? For projects in machine learning, what is 
% the training/test data that was used, and why is it realistic or
% interesting? Exactly what performance data did you collect and how 
% are you presenting and analyzing it? Comparisons to competing methods 
% that address the same problem or to variations of your own algorithm 
% are particularly useful.
\begin{itemize}
\item Each algorithm was run on 3 different initial parameter vectors on the
physical robot. We stop each run after plateauing results (no improvement 
for half the policies seen so far). We evaluate each method based on the
amount of improvement seen from the initial parameter vectors, and on the
fastest speeds achieved during runs.
\end{itemize}

@@

@@

\subsection{Discussion}
% Is your hypothesis supported? What conclusions do the results support
% about the strengths and weaknesses of your method compared to other 
% methods? How can the results be explained in terms of the underlying 
% properties of the algorithm and/or the data.

>>>>

\section{Introduction}
% Motivate and abstractly describe the problem you are addressing and
% how you are addressing it. What is the problem? Why is it important?
% What is your basic approach? A short discussion of how it fits into
% related work in the area is also desirable. Summarize the basic results
% and conclusions that you will present.
We are using machine learning to design a gait for a quadruped
robot. The robot has an on-board computer, drivers for the motors, and
will have camera feedback. The code so far consists of the robot
class, optimization class, parameterized model, sine model, and camera
feedback.  Both the hardware and software are progressing on schedule.
Some promising gaits have been generated, and we anticipate a final
evaluation of the robots gait in November.

\section{Problem definition}
% Precisely define the problem you are addressing (i.e. formally specify
% the inputs and outputs). Elaborate on why this is an interesting and
% important problem.
We are testing several different learning methods to design a gait for 
a quadruped robot from the Cornell Computational Synthesis Lab. The metric 
for evaluation of the designed gait is speed. A comparison and evaluation 
of the many different methods available for optimizing the gait of legged 
robots will be useful for future work on this challenging multidimensional 
control problem.

\section{Method}
% Describe in reasonable detail the algorithm you are using to address this 
% problem. A pseudo-code description of the algorithm you are using is 
% frequently useful. If it makes sense for your project, trace through a 
% concrete example, showing how your algorithm processes this example. The 
% example should be complex enough to illustrate all of the important aspects 
% of the problem but simple enough to be easily understood. If possible, 
% an intuitively meaningful example is better than one with meaningless 
% symbols.
We implemented and tested 8 different learning strategies:
\begin{itemize}
\item \emph{Uniform random hill climbing}: Creates a random neighbor by
  randomly choosing one parameter to adjust, and changing it completely 
  randomly in UniformStrategy(). Then evaluates this neighbor by running 
  the robot with the newly chosen parameters. If this neighbor results in a
  longer distance walked then the previous best neighbor, we save this
  neighbor as the new best neighbor. We then repeat the whole process
  using the best neighbor as a base.
\item \emph{Gaussian random hill climbing}: Creates a random neighbor by
  changing each parameter randomly based on a normal distribution. Then
  evaluates this neighbor similarly to uniform random hill climbing.
\item \emph{N-dimensional policy gradient descent}: Estimates the policy
  gradient by evaluating \emph{t} randomly generated policies
  near an initial parameter vector. Then computes the average score for
  each parameter and adjusts the base policy according to the estimated
  gradient.
\item \emph{Random}: Randomly generates policies in the range of the motion
representation.
\item \emph{Simplex (Nelder-Mead) Method}: Creates a simplex with 6 vertices and tests the fitness of each vertex. The worst vertex is reflected over the centroid. If the reflected point is better than the second worst point and worse than the best point, then the reflected point replaces the worst. If the reflected point is better than the best point, the simplex is expanded in the direction of the reflected point. The better of the reflected and the expanded point replaces the worst point. If the reflected point is worse than the second worst point, then the simplex is contracted away from the reflected point. If the contracted point is better than the reflected point, the contracted point replaces the worst point. If the contracted point is worse than the reflected point, the entire simplex is shrunk.
\item \emph{Linear regression}: To initialize, chooses five random points
and trains on them using least-squares. Then in a loop, takes a fixed-size
step in the direction of the gradient, and retrains on all points so far.
\item \emph{SVM regression}
\item \emph{Evolutionary Neural Network (HyperNEAT)\cite{2}}
\end{itemize}

@@

The quadruped robot has an on-board computer running Linux. The lower
level drivers are in C and the system is implemented in Python. Feedback about distance travelled is provided via an infrared LED and a Wii remote.

@@

\section{Experimental Evaluation}

\subsection{Methodology}
% What are the criteria you are using to evaluate your method? What
% specific hypotheses does your experiment test? Describe the 
% experimental methodology that you used. What are the dependent and 
% independent variables? For projects in machine learning, what is 
% the training/test data that was used, and why is it realistic or
% interesting? Exactly what performance data did you collect and how 
% are you presenting and analyzing it? Comparisons to competing methods 
% that address the same problem or to variations of your own algorithm 
% are particularly useful.
\begin{itemize}
\item Each algorithm was run on 3 different initial parameter vectors on the
physical robot. We stop each run after plateauing results (no improvement 
for one third of the policies seen so far). We evaluate each method based on the
amount of improvement seen from the initial parameter vectors, and on the
fastest speeds achieved during runs.
\end{itemize}

@@

\item For vector A, linear regression worked significantly better than the other algorithms, resulting in a gait (27.58 body lengths/minute) that walked over twice as fast as the next best gait (uniform random hill climbing - 11.37 body lengths/minute).

\item Vector B resulted in similarly performing gaits with all the algorithms. Random produced the best gait (17.26 body lengths/minute) and uniform random hill climbing produced the worst (9.44 body lengths/minute). The remaining algorithms each produced a gait around 13 body lengths/minute.

\item For vector C, simplex and Gaussian random hill climbing each produced a gait that substantially outperformed the other algorithms. Simplex resulted in a gait of nearly 15 body lengths/minute  and Gaussian random hill climbing produced a gait of just over 13 body lengths/minute, whereas the other algorithms returned gaits of less than 5 body lengths/minute.

\item Based on the three trials, linear regression worked the best, followed by simplex. Gaussian random hill climbing performed about as well as random, but it is unclear whether Gaussian would continue to improve and become much better than random if allowed to run for more iterations. Uniform random hill climbing and policy gradient descent performed similarly.

@@

\subsection{Discussion}
% Is your hypothesis supported? What conclusions do the results support
% about the strengths and weaknesses of your method compared to other 
% methods? How can the results be explained in terms of the underlying 
% properties of the algorithm and/or the data.
\begin{itemize}
\item  Since only three trials were tested with each algorithm and no algorithm consistently outperformed the others, there is a large standard error for each algorithm. For this reason, it is unclear if any algorithm is superior to another in this application. More trials with each algorithm would be necessary to reach a definitive ranking of the performance of the algorithms.

\item Each algorithm discovered at least one gait of over 10 body lengths/minute, including random. For this reason, we speculate that the motion model may be more critical than the learning algorithm.
\end{itemize}
